---
title: "STAT412_FINAL_PROJECT_DRAFT"
author: "Mariana Gonzalez Castro"
date: "2023-03-10"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE)
```

# CARINA AND MY SUPER AWESOME FINAL PROJECT:

```{r}
library(dplyr)
library(MASS)
```

```{r}
food_data <- read_csv("foods_dataset.csv")
data_dictionary <- read_csv("data_dictionary.csv")

```

```{r}
#Potential Response Variables
#Categorical
food_data %>% count(food_data$children)
food_data %>% count(food_data$marital_status)
food_data %>% count(food_data$employed_new)
food_data %>% count(food_data$urban)
food_data %>% count(food_data$hh_size)
food_data %>% count(food_data$edu)

food_data %>% count(food_data$super5_cat)
food_data %>% count(food_data$race)
food_data %>% count(food_data$income)

food_data %>% count(food_data$dest5_10)
```

```{r}
#children, marital_status, employed_new, urban, hh_size, edu, super5_cat, race, income, dest5_10, zip_region

food_data %>% count(food_data$zip_region) #something is up with the zip code due to the observations that have 4 digits instead of 5
```

```{r}
#Chi-square Test for Independence
#null = independent
#alt = no independent

chisq.test(food_data$children, food_data$marital_status)
chisq.test(food_data$children, food_data$employed_new)
chisq.test(food_data$children, food_data$urban) #independent
chisq.test(food_data$children, food_data$hh_size)
chisq.test(food_data$children, food_data$edu)
chisq.test(food_data$children, food_data$race)
chisq.test(food_data$children, food_data$income)
chisq.test(food_data$children, food_data$zip_region) #independent

chisq.test(food_data$marital_status, food_data$urban)
chisq.test(food_data$marital_status, food_data$hh_size)
chisq.test(food_data$marital_status, food_data$edu)
chisq.test(food_data$marital_status, food_data$race)
chisq.test(food_data$marital_status, food_data$income)
chisq.test(food_data$marital_status, food_data$zip_region)

chisq.test(food_data$urban, food_data$hh_size)
chisq.test(food_data$urban, food_data$edu)
chisq.test(food_data$urban, food_data$race)
chisq.test(food_data$urban, food_data$income)
chisq.test(food_data$urban, food_data$zip_region)

chisq.test(food_data$hh_size, food_data$edu)
chisq.test(food_data$hh_size, food_data$race)
chisq.test(food_data$hh_size, food_data$income)
chisq.test(food_data$hh_size, food_data$zip_region)

chisq.test(food_data$edu, food_data$race)
chisq.test(food_data$edu, food_data$income)
chisq.test(food_data$edu, food_data$zip_region)

chisq.test(food_data$income, food_data$race)
chisq.test(food_data$income, food_data$zip_region)

chisq.test(food_data$race, food_data$zip_region)


```

Based on Chi-square Test for independence, we may experience collinearity in the model due to the categorical variables being dependent. 


```{r}
#df = subset(food_data, select = c('children','marital_status', 'employed_new', 'urban', 'hh_size', 'edu', 'race', 'income'))
```




```{r}
#Numerical
hist(food_data$con5_10)
hist(food_data$regional_10000)
hist(food_data$dest5_10)
hist(food_data$mix5_10)
hist(food_data$design5_10)
hist(food_data$auto_bg100)
hist(food_data$poverty_tr100)

```


```{r}
#Potential Outcome Variables
hist(food_data$fruit_paid)
hist(food_data$ln_fruit_paid)

hist(food_data$vege_paid)
hist(food_data$ln_vege_paid)
```



Notes about the variables: 

Don't know wtf is up with variable **age**. It appears to be categorical ranging 1-4.

Plenty of categorical variables to work from, could do modeling based on demographics only (we had discussed modeling with only categorical variables)

The numerical variables are difficult to interpret context-wise and lowkey the article ain't that great with their explanations

I now see why there was a log transformation applied to the fruit and vegetable outcome variables. The original variables are greatly skewed. We could explore a boxcox transformation and compare? 

Types of regression models could work with: M-estimation robust (due to lot of outliers in outcome variables even after log transformation), maybe kernel? tbh not sure yet

our variables to use in our model: children + marital_status + employed_new + urban + hh_size + edu + race + income

maybe: zip_region need to talk to Carina about it

##Fruit

**Basic Linear Model**
```{r}
#start with basic linear model 
attach(food_data)
fmod0 <- lm(ln_fruit_paid ~ children + marital_status + employed_new + urban + hh_size + edu + race + income)
summary(fmod0)
plot(fmod0)
```

**Notes about the model**

*Residuals vs Fitted Plot* = appears we got constant variance, this is good

*Normal Q-Q Plot* = it's mostly normal, but the original variable has a worse distribution compared to the log transformation, we got at least 3 outliers that we gotta check 

*Residuals vs Leverage* = does not look good, got tunnel pattern...signs of collinearity (just as I  predicted)

*Summary* = all significant predictors but could that be due to collinearity? most significant are marital_status, hh_size, income

**Checking the Outliers for this linear model:**

```{r}
plot(fmod0, 4, id.n=10)

threshold <- 4/22448
cooksD <- cooks.distance(fmod0)

outlier <- cooksD[(cooksD > threshold)]
outlier

library(car)
outlierTest(fmod0)

```

##Vegetable

**Basic Linear Model**
```{r}
#start with basic linear model 
attach(food_data)
vmod0 <- lm(ln_vege_paid ~ children + marital_status + employed_new + urban + hh_size + edu + race + income)
summary(vmod0)
plot(vmod0) #gonna have similar results to the fruit model but different significant predictors
```


