---
title: "EDA"
author: "Carina Kalaydjian"
date: "2023-03-10"
output: pdf_document
---

```{r}
library(readr)
data <- read_csv("data_table1-3.csv")
View(data)
```

is log transform needed ???
```{r}
hist(data$fruit_paid, main = "fruit")

hist(data$vege_paid, main = "veg")

#####################
# now against their log transformed
hist(data$ln_fruit_paid, main = "log fruit")

hist(data$ln_vege_paid, main = "log veg")
```
log transform seems to be helpful for fruit and veg 

################################
my own data dictionary 

household_code = Numeric?, unique id numbers 
fruit_paid = Numeric
vege_paid = Numeric
ln_fruit_paid = Numeric, log of fruits
ln_vege_paid = Numeric, log of veggies
super5_cat = Categorical
con5_10 = Numeric
regional_10000 = Numeric
dest5_10 = Numeric
mix5_10 = Numeric
design5_10 = Numeric
auto_bg100 = Numeric
poverty_tr100 = Numeric
edu = Categorical 
income = Categorical
race = Categorical           
hh_size = Categorical
age = Categorical
marital_status = Categorical
children = Categorical
employed_new = Categorical
urban = categorical
msa = ? OUT
zip =  transform to be categorical of west, east, middle, south : zippcodes of households 
```{r}
library(dplyr)
# "household_code" not included 
numerical_columns = c("fruit_paid", "vege_paid", "ln_fruit_paid", "ln_vege_paid", "con5_10", "regional_10000", "dest5_10", "mix5_10", "design5_10", "auto_bg100","poverty_tr100")

categorical_columns = c("super5_cat", "edu","income", "race","hh_size","age","marital_status", "children", "employed_new", "urban", "zip")
 

i = 1
for (i in i:length(categorical_columns)){
  t = table(data[(categorical_columns[i])])
  print(t)
}


j = 1
for (j in j:length(numerical_columns)){
  name  = numerical_columns[i]
  hist(data[name], main=paste("Hist", name))
}


```
our data dosent follow normal even with log transforms 

Things to do
- explore outlines 
- relabel categories for categorical variables(except binary), to be 0, 1, 2, 3 
- explore different transformations other than log (not biggest priority, hopes and dreams)
- write code to collapse zip codes by west east south north 
- decode numerical variables a bit more 

- check for indepencene/colinearity after modeling?

Types of models 
- veggies vs everything else 
- fruit vs everything else 
- fruit or veggies vs only demographics , just cause 



Models to try 
- basic linear model 
- m-estimation rohbust in hw 1or 2? good model for data with outliers and when data dosent follow a normal distribution 
- explore more model options























